---
title: "Covid-19 death and vaccination base on time series data"
subtitle: ""
author: Yao Lu
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.


```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```



# Summary/Abstract
_Write a summary of your project._

covid vaccination data comes from
https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c
covid symptom and death data comes from
https://ga-covid19.ondemand.sas.com/docs/ga_covid_data.zip

All the raw data are saved at raw_data

Here I plan to explore these documents.

Here is the initial plan. First do ARIMA model to check the best time series model for the infection(may measured by death or symptom, haven't decided)

Second, do regression for the relationship between the covid infection and vaccination. Maybe add some other variables like population, age, gender, race. Plan to use county as individual observation. 

{{< pagebreak >}}


# Introduction 

## General Background Information
_The main data we use is the vaccination data and the data of the COVID cases in Georgia. Then after we make a summary of the data. We try some models which could guide us to do further study.Our research question is, does the coverage of vaccine decrease the spread of COVID. Under the background of the pandemic, people generate the vaccine to prevent the spread of the COVID. How many effect do the vaccine have? Now, the method we use here is using crude model to briefly test the relationship between coverage of the vaccine and incerasing speed of the COVID. _ 

## Description of data and data source
_Firstly, we load data and see if there is the need of clean and transformation_

```{r}
library(readxl)
library(readr)
library(DataExplorer)
library(dplyr)
library(ggplot2)
library(here)

data_location <- here::here("data","raw_data","Georgia_DPH_PUBLIC_Vaccination_Public_Data_in_Excel.xlsx")
vacin <- readxl::read_excel(data_location,sheet=2)

data_location1 <- here::here("data","raw_data",'ga_covid_data',"county_cases.csv")

cases <- read_csv(data_location1)

#Both vacin and cases have some unknown,so we drop them
vacin <- vacin %>% filter(COUNTY_ID!=99999)
vacin <- vacin %>% filter(COUNTY_NAME!="GEORGIA")
cases <- cases %>% filter(population!=0)

temp <- vacin[44,]
vacin[44,] <- vacin[43,]
vacin[43,] <- temp

combine <- cbind(vacin,cases)

#There are many variables which we won't need. So we clean some of them
combine1 <- combine %>% dplyr::select(COUNTY_NAME,PERSONVAX,PCTCUMPVAX, population, `14 day case rate`, `14 day cases`)
#Prepare for the two plot, one is absolute plot(None vac person vs 14 absolute number)
#the other is percentage plot (None vac percentage vs 14 day cases percentage)
combine1$novac <- combine1$population-combine1$PERSONVAX
combine1$PCTCUMPVAX <- as.numeric(combine1$PCTCUMPVAX)


complement <- function (x) {
  x <- 100-x
}

combine1$novacpct <- complement(combine1$PCTCUMPVAX)

#this are subsituted by  complement function
#for(i in 1:nrow(combine1)) {
#  combine1$novacpct[i] <- 100-as.numeric(combine1$PCTCUMPVAX[i])
#}

per1000 <- function(x){
  x <- x/1000
}

combine1$`14daycasepct` <- per1000(combine1$`14 day case rate`)

#for(i in 1:nrow(combine1)) {
#  combine1$`14daycasepct` <- combine1$`14 day case rate`[i]/1000
#}

#now we get the data we want
#View(combine1)

hist(combine1$`14daycasepct` )

```



## Questions/Hypotheses to be addressed



# Methods 




## Data aquisition



## Data import and cleaning



## Statistical analysis



{{< pagebreak >}}


# Results

## Exploratory/Descriptive analysis





## Basic statistical analysis



## Full analysis



{{< pagebreak >}}


# Discussion

## Summary and Interpretation


## Strengths and Limitations


## Conclusions



{{< pagebreak >}}

# References



