---
title: "An example exploratory analysis script"
author: "Yao Lu"
date: "03/11/2023"
output: html_document
---

  
  
This Quarto file loads the cleaned data and does some exploring.



# Setup





Load the chunk so we can pull in R script

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('exploratory.R')
```

Load the packages
```{r, packages, message = FALSE, warning = FALSE}
```

Load the data.
```{r,loaddata}
```


# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r,table}
```

# Data exploration through figures

Histogram plots for the continuous outcomes. 

'14 day case rate' first.

```{r,figurey}
```

Here we find there are two county which have high '14 day case rate'. We want to know which they are.

```{r,figurey1}
```

They are 'Quitman' and 'Stewart' counties. In the previous 14 days/100,000 population, Quitman has 392.33 cases reported, and Stewart has 1125.8 cases reported. Beyond our current research, we should pay attention about what situations make these two counties have such a high level than other counties. Previous 14 days before 01/31/2023.


Now population, RTCUMVAXADMIN,PCTCUMPVAX,PCTCUMPCVAX and PCTBOOSTER. Since last four should be highly correlated. The analysis we will apply later should be same. 
Our final model should include two predictors which are population and one of these four.

Since the four choices share same statistics steps. We just pick population+PCTBOOSTER as example. I was thinking where could I find a categorical variables. Such as could I make population as 'small','medium' and 'large'. I think we can try both way how we define population. Since there is not other good choice of categorical variable I can find in this dataset. To make the diversity of our predictors. I decide to define population as categorical variables.(However,I am not saying this a better way.)

Now 'PCTBOOSTER'.

```{r,figurex5}
```

Now '14 day case rate' as function of 'PCTBOOSTER'.

```{r,function1}
```

Here we find that two outliers make our plot hard to see clearly. We can try another model removing these two outliers. Here I just leave it here to see what our model shows.

Once more cases as function of PCTBOOSTER, stratified by population. 

```{r,function2}
```


# Notes

Here we can see both the two outliers have small population.

