---
title: "An example cleaning script with code pulled in"
author: "Yao Lu"
date: "2023-03-11"
output: html_document
---


# Processing script with code pulled in

This is essentially the same as the other Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script using the code chunk labels.


# Setup

This needs to run to load the R script so we can include the code chunks below.

Note that you can pull in more than one R script if you want.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processingcode1.R')
```




Load needed packages. Instead of having R commands here, this just pulls in the code from the R script we loaded above and labelled.


```{r, packages, message = FALSE, warning = FALSE}
```


# Data loading

Covid vaccination data comes from [here](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c).

Covid symptom and death data comes from [here](https://ga-covid19.ondemand.sas.com/docs/ga_covid_data.zip)

```{r, loaddata}
```

# Explore data
```{r, exploredata}
```
By head() of these two dataframe, we found that both cases and vacin have 161 obs. The first row in 'vacin' is GEORGIA. However, the first row in 'cases' is 'Appling'. Here I would like to combine these two dataset. So, later we can find that how much can vacin reduce the speed of COVID-19 spreading.

# Cleaning
We can find in both of these dataset, there is county_ID columns which should be identical. So, I'd like to combine these two dataset by county_ID. But they have different type. First we only get the numeric part of county_id in 'cases'.

```{r,cleandata1}
```

By check both the heads and tails of two datasets. For 'vacin', I want to remove row 1 (ID=00000), since we don't need the entire Georgia information. I want to remove row 161(ID=99999), this is an unknown county which we don't need. For 'cases', I want to remove first 2 rows which are basicly unknown  

```{r,cleandata2}
```

Then we check if all of these are matched.

```{r,cleandata3}
```

It's ready to combine by 'county ID'.

```{r,cleandata4}
```

Here we get a new dataset with 28 columns. I want to simplify that a little bit by removing the columns which are irrelevant with our research question.

```{r,cleandata5}
```

Finally, 21 columns are left. After thinking about exploring. Finally, I decide to set '14 day case rate' as my major outcome. Set 'population','RTCUMVAXADMIN','PCTCUMPVAX','PCTCUMPCVAX' and 'PCTBOOSTER' as the predictors.

#Code book

```{r,codebook}
```

Looks like the data source owner delete PCTCUMPBVAX from the data. However, haven't added 'PCTBOOSTER' to codebook. Here we assume they are same.

After we have a look at explorary part. We need to switch 'PCTCUMPVAX','PCTCUMPCVAX' and 'PCTBOOSTER' to numeric.

```{r,cleandata6}
```

Create a new categorical variables from population. 
```{r,cleandata7}
```

# Save data 

Then we save our data.
```{r,savedata}
```
